---
title: "Data Science R Basics Edx - Ariadnita la huerfanita"
output: html_document
date: "2024-10-25"
---

Installing packages:
```{r}
library(tidyverse)
library(dslabs)
```

Importing data
```{r}
data("murders")
```

Graphic of the murders and population
```{r}
murders %>% 
  ggplot(aes(population, total, label = abb, color = region)) +
  geom_label()
```



```{r}
# assigning values to variables
a <-2
b <-3
c <- -2

# solving the quadratic equation (sqrt) calcula la raiz cuadrada.
(-b + sqrt(b^2 - 4*a*c))/(2*a)
(-b - sqrt(b^2 - 4*a*c))/(2*a)
```

##It's important to remember that the functions are evaluated from in to the outside: 
```{r}
log(1) #1 es el argumento
log(exp(1)) #exponente sera el primer argumente y luego el log
```

Para ver todos los conjuntos de datos disponibles que tiene R data():
```{r}
data("murders")
```

```{r}
class(murders)
```

```{r}
str(murders)
```
```{r}
head(murders)
```
```{r}
murders$population
```

```{r}
pop <- murders$population
length(pop)
class(pop)
```
Character vectors: Si escribo a y asigno 1. Entonces si escribo a en R, me devolvera un 1.
Pero, si quiero trabajar con una cadena de caracteres a, necesito usar comillas.
```{r}
a <- 1
a
"a"
```

```{r}
class(murders$state)
```

Logical Vector:
```{r}
z <- 3 == 2 # == is a relational operator, diferente from = that will assign 2 to 3, and give us an error
z
class(z)
```

```{r}
class(murders$region) # factors are useful to restoring categorical data. Regions are categorical, there's four categories, that you can see with the function levels, Factors creates some confusion because can be taken as characters

levels(murders$region)
```

Factors creates some confusion because can be taken as characters

##Vectors
En un marco de datos como el de "murders" each column is a vector. One way to create vectors is by using the function "c" wich means concatenate (concatenar)
```{r}
codes <- c(350,222,223)
class(codes)
```
We use "" when it's a character
```{r}
country <- c("italy","canada","egypt") 
```

We can also concatenate every country wich his code
```{r}
codes <- c(italy=350, canada=222, egypt=223) #You can use "" to define the names but is exactly the same as this function
codes
```

Another function that do the same as before to concatenate countries and codes:
```{r}
names(codes) <- country
codes
```

## Function SEQ
```{r}
seq(1,10) #seq mean sequence, the first argument defines the start and the second the end of the sequence. The third arguments (1,10,1) defines the jump of the sequence
seq(1:10) #we use : if we want integers
```

## Subseting 
We use [] to access elements of a vector
```{r}
codes[2]
```
We can get more than an entrande by using
```{r}
codes[c(1,2)]#Now we get the first and second elements
codes[c(1:3)] #Can get from 1 to 3
```

```{r}
codes["canada"]
codes[c("canada","egypt")]
```

## Vector Coercion

It's an attempt of R to be flexible with the data. When an entrance doesn't work R try to guess what we are trying to do before give us an error.
But this also can make some confussion. 


```{r}
x <- c(1,"Canada",3)
x
class(x)
```
Here R has converted 1 and 3 in Character, because of "Canada".

You can converted numbers to characters with the functions as.character:
```{r}
x <- 1:5 #numbers
y <- as.character(x)
y
```

```{r}
as.numeric(y)
```

##NA
Missing data is really comun (Not available): For example
```{r}
x <- c("1","b","3")
class(x)
```
```{r}
as.numeric(x)
```

R doesn't know what to do, that's why instead of give us an error, put "b" as an NA. 


##Sorting
Sort the vector in increasing order
```{r}
sort(murders$total)
```

Order, returns the indices that sort the vector parameter
```{r}
x <- c(31, 4, 15, 92, 65)
x
```
```{r}
sort(x)
```
```{r}
index <- order(x)
x[index]
```

Again
```{r}
x
```
```{r}
order(x) #this is the index that puts the vector x in order
```

##SORTING

```{r}

sort(x)    # puts elements in order

index <- order(x)    # returns index that will put x in order
x[index]    # rearranging by this index puts elements in order
order(x)

murders$state[1:10]
murders$abb[1:10]

index <- order(murders$total)
murders$abb[index]    # order abbreviations by total murders

max(murders$total)    # highest number of total murders
i_max <- which.max(murders$total)    # which max give us the position of the highest number
murders$state[i_max]    # state name with highest number of total murders

x <- c(31, 4, 15, 92, 65)
x
rank(x)    # returns ranks (smallest to largest)
```

## Vector Arithmetic
```{r}
murders$state[which.max(murders$population)]
max(murders$population)
```

California is the state with the largest population, that's why is unfair to comparate to other states by using the murders indicator

What we have to calculate is the murders per capita

```{r}
heights <- c(69,62,66,70,70,73,67,73,67,70)
heights * 2.54 # to convert the value to centimeters
```

Para calcular la tasa de homicidios por cada 100 000, se tiene que dividir el total de homicidios entre la poblacion del mismo estado, y multiplicarlo por la tasa de 100 000 personas: 

```{r}
murders_rate <- murders$total/murders$population*100000
```

```{r}
murders$state[order(murders_rate, decreasing=TRUE)]
```

##Indexing

```{r}
index <- murders_rate < 0.71
index <- murders_rate <= 0.71
index
```

```{r}
murders$state[index]
sum(index)
```

Now we want a state that is in the montains and the rate is less than 1 x 100 000 person

```{r}
west <- murders$region == "West"
safe <- murders_rate <= 1


index = west & safe

index

murders$state[index]
```
 
## WHICH
 The index of which help us to get the true values of a sentence.
 
```{r}
index <- which(murders$state == "Massachusetts")
index
```

```{r}
murders_rate[index]
```

##MATCH

```{r}
index <- match(c("New York","Florida","Texas"),murders$state)
index
murders$state[index]
murders_rate[index]
```

%>% 




